RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-07-16 paper/RDF validation requirements/RDF validation requirements.tex
constraining elements list completed

@@

*Subsumption & $\sqsubseteq$ & sub-classes \\
*Class Equivalence & $\equiv$ & equivalent classes \\
*Sub Properties & $\sqsubseteq$ & sub-properties \\
*Property Domain & $\exists$, $\sqsubseteq$ & domain \\
*Property Range & $\sqsubseteq$, $\forall$ & range \\
*Inverse Object Properties & inverse & \\
%*Symmetric Object Properties & $\equiv$, inverse & symmetric \\
*Symmetric Object Properties & symmetric &  \\
%*Asymmetric Object Properties & $\sqsubseteq$, inverse & asymmetric \\
*Asymmetric Object Properties & asymmetric &  \\
%*Reflexive Object Properties & $\exists$, $\sqsubseteq$ & \textcolor{green}{reflexive} \\
*Reflexive Object Properties & reflexive &  \\
%*Irreflexive Object Properties & $\exists$, $\neg$, $\sqsubseteq$ & irreflexive \\
*Irreflexive Object Properties & reflexive, $\neg$ & irreflexive \\
Disjoint Properties & $\sqsubseteq$, $\neg$ & $\ne$ \\
Disjoint Classes & $\sqcap$, $\sqsubseteq$ & $\ne$ \\
Context-Sp. Property Groups & $\sqcap$ & property groups \\
Context-Sp. Inclusive OR of P. & $\sqcup$ & inclusive or \\
Context-Sp. Inclusive OR of P. Groups & $\sqcup$, $\sqcap$ & inclusive or, \\
                                                          && property groups \\
Recursive Queries & - & recursive \\
Individual Inequality & $\ne$ & \\
*Equivalent Properties & $\equiv$ & equivalent properties \\
Property Assertions & = or $\ne$ & \\
Data Property Facets & - & XSD or SPARQL constraining facets \\
                        && examples: regex, xsd:pattern, ... \\
Literal Pattern Matching & - & regex or xsd:pattern \\
Negative Literal Pattern Matching & - & regex or xsd:pattern, $\neg$ \\ 
*Object Property Paths & $\sqsubseteq$ & property paths \\
*Intersection & $\sqcap$ & intersection \\
*Disjunction & $\sqcup$ & disjunction \\
*Negation & $\neg$ & negation \\
*Existential Quantification & $\exists$ & existential quantification \\
*Universal Quantification & $\forall$ & universal quantification \\
*Minimum Unqualified Cardinality & $\geq$ & \\
*Minimum Qualified Cardinality & $\geq$ & \\
*Maximum Unqualified Cardinality & $\leq$ & \\
*Maximum Qualified Cardinality & $\leq$ & \\
*Exact Unqualified Cardinality & $\geq$, $\leq$, $\sqcap$ & = \\
*Exact Qualified Cardinality & $\geq$, $\leq$, $\sqcap$ & = \\
*Transitive Object Properties & $\sqsubseteq$ & transitive \\
Context-Sp. Exclusive OR of P. & $\neg$, $\sqcap$, $\sqcup$ & XOR \\
Context-Sp. Exclusive OR of P. Groups & $\neg$, $\sqcap$, $\sqcup$ & XOR, $\sqcap$ \\
Allowed Values & $\sqcup$ & allowed \\
Not Allowed Values & $\sqcup$, $\neg$ & not allowed \\
%Literal Ranges & $\geq$, $\leq$, $\sqcap$ &  \\ 
Literal Ranges & - & xsd:minInclusive \\
               & & xsd:maxInclusive \\
%Negative Literal Ranges & $\geq$, $\leq$, $\sqcap$, $\neg$ & \\  
Negative Literal Ranges & - & xsd:minInclusive \\
                        & & xsd:maxInclusive \\
Required Properties & $\exists$ & required \\

@@

Optional Properties & $\exists$, $\sqsubseteq$ & optional properties \\
Repeatable Properties & $\geq$ & repeatable properties \\
Negative Property Constraints & $\neg$, $\exists$ & negative properties \\
*Individual Equality & = & \\
*Functional Properties & functional & \\ 
*Inverse-Functional Properties & inverse, functional & inverse functional \\
*Value Restrictions & $\exists$ & value restriction \\
*Self Restrictions & $\exists$ & self restriction \\
Primary Key Properties & inverse, functional & keyfor or primary key \\
*Class-Specific Property Range & $\sqsubseteq$, $\exists$, $\neg$ & range \\
*Class-Sp. Reflexive Object P. & reflexive & \\
Membership in Controlled Vocabularies & $\forall$, $\sqcap$, $\sqcup$ & membership in \\
  && controlled vocabularies \\
IRI Pattern Matching & - & IRI pattern matching \\
Literal Value Comparison & $>$ or $\geq$ or $<$ or & \\
                         & $\leq$ or = or $\ne$ & \\
Ordering & - & order \\ 
Validation Levels & - & validation level \\ 
String Operations & - & SPARQL string operations \\ 
  && examples: STRLEN, \\ 
	&& SUBSTR, CONTAINS \\
Context-Specific Valid Classes & - & context-specific valid classes \\
Context-Specific Valid Properties & - & context-specific valid properties \\
Default Values & - & default values \\
Mathematical Operations & - & example: multiplication \\
Language Tag Matching & - & language tag matching \\
Language Tag Cardinality & - & language tag minimum cardinality, \\
                            && language tag maximum cardinality, \\
													  && language tag exact cardinality \\
Whitespace Handling & - & whitespace \\
HTML Handling & - & html \\
Conditional Properties & $\sqsubseteq$ & conditional \\
Recommended Properties & - & recommended \\
Handle RDF Collections & - & actions on RDF collections \\
Value is Valid for Datatype & - & value is valid for datatype \\
Use Sub-Super Relations in Validation & - & not redundant \\
*Cardinality Shortcuts & & optional \& non-repeatable, \\
                       & & optional \& repeatable \\
											 & & mandatory \& non-repeatable \\
											 & & mandatory \& repeatable \\
Aggregations & - & aggregations \\
                 && e.g., count \\
*Class-Specific Irreflexive Object Properties & reflexive, $\neg$ & irreflexive \\
Provenance & - & provenance \\   
Data Model Consistency & - & SPARQL \\    
Structure & - & SPARQL \\  
Labeling and Documentation & - & SPARQL \\   
Vocabulary & - & vocabulary \\  
HTTP URI Scheme Violation & - & HTTP URI scheme violation \\

>>>>

Property \emph{p} should be ordered, i.e., point to a list (rdf:List) of values.

\begin{gcotable}
property & $\top$ & p & - & - & ordered property & - \\
\end{gcotable}

@@

*Subsumption & $\sqsubseteq$ & sub-class \\
*Class Equivalence & $\equiv$ & equivalent class \\
*Sub Properties & $\sqsubseteq$ & sub-property \\
*Property Domains & $\exists$, $\sqsubseteq$ & property domain \\
*Property Ranges & $\sqsubseteq$, $\forall$ & property range \\
*Inverse Object Properties & inverse & inverse property \\
%*Symmetric Object Properties & $\equiv$, inverse & symmetric \\
*Symmetric Object Properties & symmetric & symmetric property \\
%*Asymmetric Object Properties & $\sqsubseteq$, inverse & asymmetric \\
*Asymmetric Object Properties & asymmetric & asymmetric property \\
%*Reflexive Object Properties & $\exists$, $\sqsubseteq$ & \textcolor{green}{reflexive} \\
*Reflexive Object Properties & reflexive & reflexive property \\
%*Irreflexive Object Properties & $\exists$, $\neg$, $\sqsubseteq$ & irreflexive \\
*Irreflexive Object Properties & reflexive, $\neg$ & irreflexive property \\
Disjoint Properties & $\sqsubseteq$, $\neg$ & disjoint property \\
Disjoint Classes & $\sqcap$, $\sqsubseteq$ & disjoint class \\
Context-Sp. Property Groups & $\sqcap$ & property group \\
Context-Sp. Inclusive OR of P. & $\sqcup$ & inclusive or \\
Context-Sp. Inclusive OR of P. Groups & $\sqcup$, $\sqcap$ & inclusive or \\
Recursive Queries & - & recursive \\
Individual Inequality & $\ne$ & different individuals \\
*Equivalent Properties & $\equiv$ & equivalent properties \\
Property Assertions & = or $\ne$ & =, $\ne$ \\
Data Property Facets & - & SPARQL functions: \\
                        && REGEX, STRLEN \\
												&& XML Schema constraining facets: \\
                        && xsd:length, xsd:minLength, xsd:maxLength, xsd:pattern, \\
												&& xsd:enumeration, xsd:whiteSpace, xsd:maxInclusive, \\
												&& xsd:maxExclusive,xsd:minExclusive, xsd:minInclusive, \\
												&& xsd:totalDigits, xsd:fractionDigits \\
Literal Pattern Matching & - & REGEX, xsd:pattern \\
Negative Literal Pattern Matching & - & REGEX, xsd:pattern and $\neg$ \\ 
*Object Property Paths & $\sqsubseteq$ & property path \\
*Intersection & $\sqcap$ & intersection \\
*Disjunction & $\sqcup$ & disjunction \\
*Negation & $\neg$ & negation \\
*Existential Quantification & $\exists$ & existential quantification \\
*Universal Quantification & $\forall$ & universal quantification \\
*Minimum Unqualified Cardinality & $\geq$ & $\geq$ \\
*Minimum Qualified Cardinality & $\geq$ & $\geq$ \\
*Maximum Unqualified Cardinality & $\leq$ & $\leq$ \\
*Maximum Qualified Cardinality & $\leq$ & $\leq$ \\
*Exact Unqualified Cardinality & $\geq$, $\leq$, $\sqcap$ & = \\
*Exact Qualified Cardinality & $\geq$, $\leq$, $\sqcap$ & = \\
*Transitive Object Properties & $\sqsubseteq$ & transitive property \\
Context-Sp. Exclusive OR of P. & $\neg$, $\sqcap$, $\sqcup$ & exclusive or \\
Context-Sp. Exclusive OR of P. Groups & $\neg$, $\sqcap$, $\sqcup$ & exclusive or \\
Allowed Values & $\sqcup$ & allowed values \\
Not Allowed Values & $\sqcup$, $\neg$ & not allowed values \\
%Literal Ranges & $\geq$, $\leq$, $\sqcap$ &  \\ 
Literal Ranges & - & xsd:minInclusive, xsd:maxExclusive \\
               & & xsd:maxInclusive, xsd:minExclusive \\
%Negative Literal Ranges & $\geq$, $\leq$, $\sqcap$, $\neg$ & \\  
Negative Literal Ranges & - & $\neg$ and \\  
                        && xsd:minInclusive, xsd:maxExclusive \\
                        && xsd:maxInclusive, xsd:minExclusive \\
Required Properties & $\exists$ & required property \\

@@

Optional Properties & $\exists$, $\sqsubseteq$ & optional property \\
Repeatable Properties & $\geq$ & repeatable property \\
Negative Property Constraints & $\neg$, $\exists$ & negative properties \\
*Individual Equality & = & equal individuals \\
*Functional Properties & functional & functional property \\ 
*Inverse-Functional Properties & inverse, functional & inverse functional property \\
*Value Restrictions & $\exists$ & value restriction \\
*Self Restrictions & $\exists$ & self restriction \\
Primary Key Properties & inverse, functional & primary key \\
*Class-Specific Property Range & $\sqsubseteq$, $\exists$, $\neg$ & property range \\
*Class-Sp. Reflexive Object P. & reflexive & reflexive property \\
Membership in Controlled Vocabularies & $\forall$, $\sqcap$, $\sqcup$ & membership in controlled vocabularies \\
IRI Pattern Matching & - & IRI pattern matching \\
Literal Value Comparison & $>$ or $\geq$ or $<$ or & $>$, $\geq$, $<$, $\leq$, =, $\ne$ \\
                         & $\leq$ or = or $\ne$ & \\
Ordering & - & ordered properties, ordered values \\ 
Validation Levels & - & validation level \\ 
String Operations & - & SPARQL string functions: \\ 
                     && STRLEN, SUBSTR, UCASE, LCASE, \\
										 && STRSTARTS, STRENDS, CONTAINS, \\ 
										 && STRBEFORE, STRAFTER, ENCODE\_FOR\_URI, \\
										 && CONCAT, langMatches, REGEX, REPLACE \\
Context-Specific Valid Classes & - & context-specific valid classes \\
Context-Specific Valid Properties & - & context-specific valid properties \\
Default Values & - & default values \\
Mathematical Operations & - & addition, subtraction, multiplication, division \\
Language Tag Matching & - & language tag matching \\
Language Tag Cardinality & - & language tag minimum cardinality, \\
                            && language tag maximum cardinality, \\
													  && language tag exact cardinality \\
Whitespace Handling & - & no whitespaces \\
HTML Handling & - & no html \\
Conditional Properties & $\sqsubseteq$ & conditional property \\
Recommended Properties & - & recommended property \\
Handle RDF Collections & - & actions on RDF collections: \\
                          && append list element \\
Value is Valid for Datatype & - & value is valid for datatype \\
Use Sub-Super Relations in Validation & - & non redundant properties \\
*Cardinality Shortcuts & & optional \& non-repeatable property, \\
                       & & optional \& repeatable property, \\
											 & & mandatory \& non-repeatable property, \\
											 & & mandatory \& repeatable property \\
Aggregations & - & aggregations: \\
                 && count \\
*Class-Specific Irreflexive Object Properties & reflexive, $\neg$ & irreflexive property \\
Provenance & - & provenance information required \\   
Data Model Consistency & - & $<$SPARQL query$>$ \\    
Structure & - & $<$SPARQL query$>$ \\  
Labeling and Documentation & - & $<$SPARQL query$>$ \\   
Vocabulary & - & vocabulary \\  
HTTP URI Scheme Violation & - & HTTP URI scheme violation \\
